### 2020-10-16 README.Rmd
RUns great on SCBI!
I feel good about the readme now. Did some editing of the text. Looks great to me!

We present an example analysis using:

* [Michigan Big Woods](https://doi.org/10.7302/wx55-kt18) research plot data
* [Smithsonian Conservation Biology Institute (SCBI) ForestGEO](https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census) research plot data

@@

The data for the Big Woods plor are included in the package. For SCBI, we load and preprocess the data corresponding to the two census reflecting the time period over which we consider growth in dbh. We load them into R as "tibble" data frames thereby ensuring a standardized input/output format that can be used across all `tidyverse` packages @tidyverse. Furthermore, we ensure that the different variables have the correct names, types (`dbl`, `data`, `factor`).

@@

For SCBI we will just look at a 9 ha plot to speed calculations.

@@

We then combine the two census data frames into a single data frame  that now includes a numerical variable `growth` reflecting the average annual growth of dbh in cm. Furthermore, variables that (in theory) remain unchanged between censuses appear only once, such as `gx`, `gy`, species-related variables. Variables that should change between censuses are tagged with `1/2` indicating earlier/later, such as `dbh1/dbh2`, `codes1/codes2`.

@@

# we need to filter out the resprouts
bw_census_2014 <- bw_census_2014 %>% 
  filter(!str_detect(codes, 'R'))

@@

scbi_growth_df <- 
  # Merge both censuses and compute growth:
  compute_growth(census_df1, census_df2, id) %>%
  # they are mesuaring in mm while we are measuring in cm!!!
  mutate(growth = growth/10,
             sp = as.factor(sp))

@@

1. In order to control for study region "edge effects" (cite Waller?), we need a function that adds "buffers" of trees. In our case, since our model of interspecies competition relies on a spatial definition of who is a "neighboring competitor" and certain explanatory varibles such as biomass are cummulative, we need to ensure that all trees being modeled are not biased to have different neighbor structure, in particular the trees at the boundary of study regions.
1. Assign each tree to a "fold" for cross-validation purposes. Conventional cross-validation schemes assign observations to folds by resampling individual observations at random. However, underlying this scheme is an assumption that the observations are independent. In the case of forest census data, observations exhibit spatial autocorrelation, and thus this dependence must be incorporated in our resampling scheme @roberts2017 @pohjankukka2017. Packages that have implemented spatial cross-validation include @valavi2019

Thus, before we can incorporate the above information to `growth_df`, we need to define two constants:

@@

@@

For a focal tree of interest, the `max_dist` of `r max_dist` meters acts as a radius defining a neighborhood within which all trees are considered competitors. In other words, all trees within `r max_dist` meters of the focal tree are considered competitor trees. Other studies have estimated this distance; we used `r max_dist` meters as an average of estimated values [@canham2004, @canham2006, @uriarte2004, @tatsumi2013]. Following Tobler's first law of geography that "everything is related to everything else, but near things are more related than distant things." @tobler1970firstlawofgeo, we assume that the degree of spatial autocorrelation is inversely-related to distance. However, we further assume that once trees are more than `r max_dist` meters apart, this autocorrelation is negligeable.

We use the function `add_buffer_variable` to identifiy which trees are inside of a buffer region of size `max_dist`. The user will need to specify a study region and convert to a `sf_polygon`.

@@

We use the [`blockCV`](https://github.com/rvalavi/blockCV) package to define the spatial grid, whose elements will act as the folds in our leave-one-out (by "one" we meen "one grid block") cross-validation scheme.

@@

This requires the `growth_df` data frame; `max_dist`, the scalar defining competitive range; `cv_fold_size`, defining the size of the spatial cross-validation blocks; and the `id` variable as inputs.

@@

Now we are ready to fit the competition model with `fit_bayesian_model`. This function needs only the `focal_vs_comp` as an input. Other options allow the user to specify prior parameters and run a species identity shuffle (see below).

@@

This output has the posterior parameters for the specified competition model. This `posterior_param` output can be used to get predicted growths for each individual (with `predict_bayesain_model`) to test how well the model performs. Or this `posterior_param` output can be plots (either the betas or lambdas) to understand what controls individual growth.

@@

Now we test whether the identity of the competitor matters. We do this by shuffling the identity of competitors (but not of focal trees or spatial locations or sizes) and fitting the model again. We then compare RMSEs to see whether competitor identity matters to competitive effects

@@

The RMSE is lower for the non-shuffled version. This gives support for the idea that competitor identiy does matter for competitive interactions.

@@

@@

@@

For the above results we fit the model to the entire data set, and then make predictions across the entire data set from that fit. This could lead to overfitting because we are using the training data to also test the model. If model error is spatially correlated this could be a large issue (cite important sources here!). We can use the spatial block structure we defined above to deal with with. The function `run_cv` goes through each fold in the `cv_grid` and fits the model on all the other folds. Then applies that fit to the focal fold. It is a wrapper for `fit_bayesain_model` and `predict_bayesain_model` but fits a separate model for each fold.

@@

@@

We might be interested in the posterior distributions of parameters. The betas tell us about how fast each species grows and how this depends on DBH. The lambdas, which are often of more interest, are the species-specific competition coefficents The full lambda matrix gives competition strength between species. There is a rich literature how this matrix (cite).

Because of the structure of the `bw_fit_model` object we cannot simply draw these curves based on the posterior distribution. `bw_fit_model` gives the parameteres *compared* to a baseline. This is often not of interest. So to display these parameters as we care about them we have to sample from the baseline distribution and from the comparison one to get the posterior distribution of interest.

>>>>

We present an example analysis using data from two [ForestGEO](https://forestgeo.si.edu/) plots:

* [Michigan Big Woods](https://doi.org/10.7302/wx55-kt18) research plot data. For more information see [here](http://hdl.handle.net/2027.42/156251).
* [Smithsonian Conservation Biology Institute (SCBI) ForestGEO](https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census) research plot data. For more information see [here](https://doi.org/10.1890/13-0010.1).

@@

The data for the Big Woods plot are included in the package. For SCBI, we load and preprocess the data corresponding to the two census reflecting the time period over which we consider growth in dbh. We load them into R as "tibble" data frames thereby ensuring a standardized input/output format that can be used across all `tidyverse` packages @tidyverse. Furthermore, we ensure that the different variables have the correct names, types (`dbl`, `data`, `factor`).

@@

For SCBI we will look at just a 9 ha section of the plot to speed calculations.

@@

For each plot we combine the two census data frames into a single data frame  that now includes a numerical variable `growth` reflecting the average annual growth of dbh in cm. Furthermore, variables that (in theory) remain unchanged between censuses appear only once, such as `gx`, `gy`, species-related variables. Variables that should change between censuses are tagged with `1/2` indicating earlier/later, such as `dbh1/dbh2`, `codes1/codes2`.

@@

# we need to filter out the re-sprouts
bw_census_2014 <- bw_census_2014 %>% 
  filter(!str_detect(codes, 'R'))

@@

scbi_growth_df <- 
  # Merge both censuses and compute growth:
  compute_growth(census_df1, census_df2, id) %>%
  # Convert growth from cm to mm to make result comperable
  mutate(growth = growth/10,
             sp = as.factor(sp))

@@

1. In order to control for study region "edge effects" (cite Waller?), we need a function that adds "buffers" of trees. In our case, since our model of interspecific competition relies on a spatial definition of who is a "neighboring competitor" and certain explanatory variables such as biomass are cumulative, we need to ensure that all trees being modeled are not biased to have different neighbor structure, in particular the trees at the boundary of study regions.
1. Assign each tree to a "fold" for cross-validation purposes. Conventional cross-validation schemes assign observations to folds by resampling individual observations at random. However, underlying this scheme is an assumption that the observations are independent. In the case of forest census data, observations exhibit spatial autocorrelation, and thus this dependence must be incorporated in our resampling scheme @roberts2017 @pohjankukka2017. Packages that have implemented spatial cross-validation include @valavi2019

Thus, before we can incorporate the above information to `growth_df`, we need to define two constants.

@@

The first, `cv_fold_size` is the length of the folds, each fold is a square. The next, `max_dist` is the maximum distance for a tree's competitive neighborhood. Trees within this distance of each other are assumed to compete while those farther than this distance apart do not. Both these values are in the same units as `gx` and `gz`, most often this will be meters.

<!--
DA comment: I am not sure about the below so I took it out. We don't assume that competitive interactions are inversely related to distance. We assume that competitive interactions are equivalent between distance 0 and 7.5, but than zero after that. Canham has the inversely related structure but we don't.

Following Tobler's first law of geography that "everything is related to everything else, but near things are more related than distant things." @tobler1970firstlawofgeo, we assume that the degree of spatial autocorrelation is inversely-related to distance. However, we further assume that once trees are more than `r max_dist` meters apart, this autocorrelation is negligeable. -->

Other studies have estimated the value of `max_dist`; we use an average of estimated values [@canham2004, @canham2006, @uriarte2004, @tatsumi2013].

@@

We use the function `add_buffer_variable` to identify which trees are inside of a buffer region of size `max_dist`. The user will need to specify a study region and convert to a `sf_polygon`.

@@

We use the [`blockCV`](https://github.com/rvalavi/blockCV) package to define the spatial grid, whose elements will act as the folds in our leave-one-out (by "one" we mean "one grid block") cross-validation scheme. The upshot here is we add `foldID` to `growth_df` which identifies which fold each individual is in, and the creation of a `cv_grid_sf` object which gives the geometry of the cross validation grid.

@@

This requires the `growth_df` data frame; `max_dist`, the scalar defining competitive range; `cv_grid_sf`, giving the cross validation grid; and the `id` variable as inputs.

@@

Now we are ready to fit the competition model with `fit_bayesian_model`. This function needs only the `focal_vs_comp` as an input. Other options allow the user to specify prior parameters and run a competitor-species identity shuffle (see below).

@@

This output has the posterior parameters for the specified competition model. This `posterior_param` output can be used to get predicted growths for each individual (with `predict_bayesain_model`) to test how well the model performs. Or this `posterior_param` output can be plotted (either the betas or lambdas with `plot_posterior_params`) to understand what controls growth and the nature of the competitive ineractions.

@@

Now we test whether the identity of the competitor matters. We do this by shuffling the identity of competitors (but not of focal trees or spatial locations or sizes) and fitting the model again. We then compare RMSEs to see whether competitor identity matters to competitive interactions.

@@

The RMSE is lower for the non-shuffled version. This gives support for the idea that competitor identity does matter for competitive interactions.

@@

Here we provide examples of seeing how well the model fits and then a spatial plot of the residuals to see whether there are any consistent spatial patterns to them.

@@

scbi_growth_df_noCV %>%
  rmse(truth = growth, estimate = growth_hat) %>%
  pull(.estimate)

@@

For the above results we fit the model to the entire data set, and then make predictions across the entire data set from that fit. This could lead to overfitting because we are using the training data to also test the model. If model error is spatially correlated this could be a large issue (cite important sources here!). We can use the spatial block structure we defined above to deal with with. The function `run_cv` goes through each fold in the `cv_grid_sf` and fits the model on all the other folds. Then applies that fit to the focal fold. It is a wrapper for `fit_bayesain_model` and `predict_bayesain_model` but fits a separate model for each fold.

@@

Here also this RMSE is much higher than that for the above SCBI model fit without cross validation.

@@

We might be interested in the posterior distributions of parameters. The betas tell us about how fast each species grows and how this depends on DBH. The lambdas, often of more interest, are the species-specific competition coefficients. The full lambda matrix gives competition strength between species. There is a rich literature on this matrix (cite).

Because of the structure of the `bw_fit_model` object we cannot simply draw these curves based on the posterior distribution. `bw_fit_model` gives the parameters *compared* to a baseline. This is not of direct interest. So to display these parameters, as we care about them, we have to sample from the baseline distribution and from the comparison one to get the posterior distribution of interest.
